---
title: "Assigment - kNN DIY"
author:
  - Dewi Joanne - Author
  - Ted van Deelen - Reviewer
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_notebook:
    toc: true
    toc_depth: 2
---
#Setup

```{r}
library(googlesheets4)
library(class)
library(caret)

require(class)
```

###Note Ted: I would put tidyverse in this list as well. That way you have all packages together.
---

Choose a suitable dataset from [this](https://github.com/HAN-M3DM-Data-Mining/assignments/tree/master/datasets) folder and train  your own kNN model. Follow all the steps from the CRISP-DM model.


## Business Understanding

This project will perform a classification K-nearest Neighbours (KNN) on an Occupancy Detection dataset for classifying the occupancy.

The k-nearest neighbours’ algorithm (k-NN) is a non-parametric method used for classification and regression [1].

In k-NN classification, the output is a class membership. An object is classified by a plurality vote of its neighbours, with the object being assigned to the class most common among its k nearest neighbours (k is a positive integer, typically small). If k = 1, then the object is simply assigned to the class of that single nearest neighbour.

[KNN-occupancy.csv](https://archive.ics.uci.edu/ml/datasets/Occupancy+Detection+)

The accuracy of the prediction of occupancy in an office room using data from light, temperature, humidity and CO2 sensors has been evaluated with different statistical classification models using the open source program R. Three data sets were used in this work, one for training, and two for testing the models considering the office door opened and closed during occupancy. 



```{r}

```

###Note Ted: I recommend removing the chunk above since it has no code.

## Data Understanding
The data we will be using comes from the University of Wisconsin and is available online as an open source dataset https://archive.ics.uci.edu/ml/datasets/Occupancy+Detection+ (UCI Machine Learning Repository: Occupancy Detection Data Set). 

Experimental data used for binary classification (room occupancy) from Temperature,Humidity,Light and CO2. Ground-truth occupancy was obtained from time stamped pictures that were taken every minute.


Attribute Information:

date time year-month-day hour:minute:second
Temperature, in Celsius
Relative Humidity, %
Light, in Lux
CO2, in ppm
Humidity Ratio, Derived quantity from temperature and relative humidity, in kgwater-vapor/kg-air
Occupancy, 0 or 1, 0 for not occupied, 1 for occupied status

```{r}
library(tidyverse)
url <- "https://raw.githubusercontent.com/HAN-M3DM-Data-Mining/data-mining-s2y2122-dewijoanne/master/datasets/KNN-occupancy.csv"
occupancy_data <- (url)
View(occ)
```

###Note Ted: I would remove tidyverse from here and move it further up in the list. See note at line 22.
###Note Ted: Line 67 will not actually read the code. To do this, I recommend writing the code as followed: occupancy_data <- read_csv(url)
###Note Ted: 'view(occ)' does not work. I recommend changing this to 'view(occupancy_data)' which will make the view function work. As an alternative you can also just click on the data in the 'Environment' tab in the right side of your screen.

Using the str() function we can have some basic information about the dataset.

```{r}
str(occupancy_data)
```
The dataset has 7 variables (columns) and 8,143 observations (rows).

```{r}
names(occupancy_data)
head(occupancy_data)
tail(occupancy_data)
summary(occupancy_data)
```


## Data Preparation

The first variable, id, contains unique patient IDs. The IDs do not contain any relevant information for making predictions, so we will delete it from the dataset.

```{r}
new_occupancy_data <- occupancy_data[1]
head(new_occupancy_data)
```

###Note Ted: To make this code work correctly I recommend changing the '1' on line 95 to '-1'. The code right now would load in the column by itself and remove the other columns while you want to keep the entire data set except for this column.

The variable named occupancy contains the outcomes we would like to predict - ‘1’ for ‘Occupied’ and ‘0’ for ‘Not Occupied.’ The variable we would like to predict is called the ‘label.’ We can look at the counts and proportions for both outcomes, using the tables() and prop.tables()functions.

```{r}
countOccu <- table(new_occupancy_data$Occupanc)
propOccu <- round(prop.table(countOccu) * 100, digits = 0)

countOccu
propOccu
```

###Note Ted: There is a skipped letter on line 104. The final word should be changed from 'Occupanc' to 'Occupancy' to make sure the code works properly.

The variable is now coded as a type character. Many models require that the label is of type factor. This is easily solved using the factor() function.

```{r}
new_occupancy_data$Occupancy <- factor(new_occupancy_data$Occupancy, levels = c("1", "0"), labels = c("Occupied", "Not Occupied")) %>% relevel ("1")
head(new_occupancy_data, 10)
```

###Note Ted: To be able to make this code run properly, I recommend changing the last '1' of line 116 into '"Not Occupied"'. This is because '1' is no longer a value since this is now changed to 'Occupied' or 'Not occupied'.

The features consist of different situation for measuerments. We will take the first three and have a closer look.

```{r}
summary(new_occupancy_data[c("Temperature", "Humidity", "Light")])
```
The three variables have very different ranges, This could potentially cause problems for modeling. To solve this we’ll apply normalization to rescale all features to a standard range of values.

```{r}
normalize <- function(x){return((x - min(x))/ (max(x) - min(x)))}

nCols <- dim(new_occupancy_data)[6]
cleanDF_n <- sapply(1:5,
                    function(x) {
  normalize(new_occupancy_data[,x])
}) %>% as.data.frame()

summary(cleanDF_n)
```

###Note Ted: After you defined the normalize function, you could run some test sets to see if the function works properly.
###Note Ted: To make the code run well in the final stages, I recommend changing the '6' in row 132 to '2'. Otherwise the wrong column will be used which causes problems in future chunks.
###Note Ted: Personally with more complicated chunks like these, I like to split them up so that they are better readable and can be used separatley when there is an error.

When we take the variables we selected earlier and look at the summary parameters again, we’ll see that the normalization was successful.

We can now split our data into training and test sets.

```{r}
train_df <- cleanDF_n[1:6526]
testd_df <- cleanDF[6527:8143]
```

###Note Ted: I recommend using the same dataset for these 2 new datasets. In this case I recommend the use of the 'cleanDF_n' dataset since 'cleanDF' does not exist. Therefore the code wouldn't work like this.
###Note Ted: The formula does not work unless you 'specify' the columns. to take all, I suggest making the formula as followed: cleanDF_N[NumberX:NumberY, ].
###Note Ted: There is a typo in line 151. I recommend changing the name of the created dataset from 'testd_df' to 'test_df' to avoid confusion in further chunks.

When creating the training and test sets, we’ve excluded the labels. We’ll create separate training and tests sets for them too.

```{r}
train_label <- occ[1:6526, ]
test_label <- new_occupancy_data[6527:8143, ]
```

###Note Ted: In line 161 the same mistake is made as in line 68. 'occ' is not defined. I would therefore change this to 'new_occupancy_data'.
###Note Ted: To make this chunk run as it is supposed to I would add a '6' after the comma in both lines of this chunk.

## Modeling
To train the KNN model we only need one single function from the class package. It takes the set with training features and the set with training label. The trained model is applied to the set with test features and the function gives back a set of predictions.

```{r}
occupancy_test_pred <- knn(train = as.matrix(train_d), test = as.matrix(test_df), cl = as.matrix(train_label), k = 21)
head(occupancy_test_pred)
```

###Note Ted: This chunk won't work because in the first row it says 'train_d' instead of 'train_df'

## Evaluation and Deployment
Now that we have a set of predicted labels we can compare these with the actual labels. A diffusion table shows how well the model performed.


```{r}
confusionMatrix(occupancy_test_pred, test_label[[1]], positive = NULL, dnn = c("Prediction", "True"))
```